
% MO problems
\begin{frame}
\frametitle{Multi-objective problems}
\textbf{Multi-objective problems} are optimization problems involving optimizing multiple criteria.
\pause
\begin{itemize}
  \item{ Usually conflicting criterias; }\pause
  \item{ A set of \textit{optimal} (non-dominated) solutions; } \pause
  \item{ Models many real applications. }
\end{itemize}
\vfill
\end{frame}

% MOKP
\begin{frame}
\frametitle{The multi-dimensional knapsack problem}
The \textbf{multi-dimensional knapsack problem} (MOKP)
is one of the most important multi-objective problems.
\pause
\begin{itemize}
  \item{ Is a generalization of the $0-1$ knapsack problem (\nphard); }\pause
  \item{ Models project selection, capital budgeting, cargo loading, flow shop scheduling and others;} \pause
  \item{ Models many real applications; }
\end{itemize}
\end{frame}

% Motivation
\begin{frame}
\frametitle{Motivation}
Several exact approaches have been proposed in the literature for
solving the MOKP. \\ \pause
\vfill
However no effective exact method for large MOKP, expecially with
more than two objectives. \\ \pause
\vfill
Even for the bi-objective case, some medium sized instances
has shown to be hard to solve exactly. \\ \pause
\vfill
This reason motivates the development of heuristic methods.
\vfill
\end{frame}

% The Proposal
\begin{frame}
\frametitle{Proposal}
This work proposes the development of a heuristic for the MOKP
based on an evolutionary algorithm called shuffled complex evolution (SCE).
\\ \medskip \pause
As a performance improvement for the approach,
a multi-dimensional indexing strategy
will be used for handling the large amount of solutions.
\\ \medskip \pause
The SCE has been successfully used for solving optimization problems, among then,
the multi-dimensional knapsack problem (MKP) (also a contribution of this work).
\end{frame}

% Contributions and Publications
\begin{frame}
\frametitle{Contributions and Publications}\pause
\begin{enumerate}
\item{
\textbf{Efficient indexing strategy for MOKP solutions:} \pause
\begin{itemize}
 \item[{\tiny$\bullet$}] { BARONI, M. D. V; VAREJ\~AO, F. M. Multi-dimensional indexing on dynamic programming for multi-objective knapsack problem. \textit{International Transactions in Operational Research}, Wiley Online Library, 2017, Submitted. }
\end{itemize}
} \pause
\medskip
\item{
\textbf{A SCE algorithm for the MKP:} \pause
\begin{itemize}
  \item[{\tiny$\bullet$}] { BARONI, M. D. V; VAREJ\~AO, F. M. A shuffled complex evolution algorithm
  for the multidimensional knapsack problem. In: \textit{Progress in Pattern Recognition, Image Analysis, Computer Vision, and Applications.} [S.l.]: Springer, 2015. p. 768-775. }
  \pause
 \item[{\tiny$\bullet$}] { BARONI, M. D. V; VAREJ\~AO, F. M. A shuffled complex evolution algorithm
  for the multidimensional knapsack problem using core concept. In: IEEE. \textit{Evolutionary Computation (CEC), 2016 IEEE Congress on.} [S.l.], 2016 p. 2718-2723. }
\end{itemize}
} \pause
\medskip
\item{
\textbf{Hybrid heuristic for MOKP:} \pause
\begin{itemize}
 \item[{\tiny$\bullet$}] { BARONI, M. D. V; VAREJ\~AO, F. M. An efficient shuffled complex evolution algorithm for the multi-objective knapsack problem. In: IEEE \textit{Evolutionary Computation (CEC), 2018 IEEE Congress on.} [S.l.], 2018, In Preparation }
\end{itemize}
}
\end{enumerate}
\end{frame}
