
The core concept was first presented for the one-dimensional 0-1 knapsack problem (KP),
leading to very successful KP algorithms.
The main idea is to reduce the original problem by only considering a set of
items for which it is hard to decide if they will occur or not in an optimal solution.
This set of items is named core.
The variables for all items outside the core are fixed to certain values.

\begin{equation}
  e_j = \frac{p_j}{w_j},
\end{equation}

The knapsack problem considers items $j = 1, \ldots, n$, associated profits $p_j$ and
associated weights $w_j$.
A subset of these items has to be selected and packed into a knapsack having capacity $c$.
The total profit of the items in the knapsack has to be maximized, while the
total weight is not allowed to exceed $c$.
The KP Core problem (KPC) is defined as
\begin{align}
  \text{maximize} & \sum_{j \in C} p_j x_j  + \tilde{p}\\
  \text{subject to} & \sum_{j \in C} w_{j} x_j \leqslant c - \tilde{w}\\
  & x_j \in \{0, 1\}, \quad j \in C.
\end{align}

The first row represents the efficiency value of each item and the second row
represents the value of each variable on the LP-relaxation optimal solution.
The items are sorted in descending order of efficiency value.
The last row illustrates the variable fixing after the defined core.
Asterisks indicate free variables associated to the items in the
approximate core.

The previous definition of the core for KP can be extended to MKP without major
difficulties, once an efficiency measure is defined for the MKP,
as addressed in, even though,
a proper efficiency measure for MKP is not obvious due the
multidimensional weight of the items.
A well accepted efficiency measure is discussed in Section~\ref{subsec:dual}.

Now let $x^*$ be an optimal solution for a MKP and assume that the items are
sorted in descending order after some given efficiency measure. Then let
\begin{align}
  a = \min \{ j | x_j^* = 0 \}, \quad b = \max \{ j | x_j^* = 1 \}.
\end{align}

In contrast to KP, the solution of the LP-relaxation of MKP in general does not
consists of a single fractional split item.
